\TheTitle

### 2020-05-19 Paper1/ResponseToReviewers/Response_to_Reviewer.tex

\end{align*}
Since only the last term $\bm{H}^{-1}(\bm\lambda) \bm{\hat u}$ depends on the moment vector $\bm{\hat u}$, we have
\begin{align*}
\partial_{\bm{\hat u}} \bm d(\bm{\lambda},\bm{\hat u}) = \bm{H}^{-1}(\bm\lambda).

@@

\\ \textbf{Response:} We added a short description on how the defined quantities relate to the conserved variables TODO.
\item I missed the definition of the reference solution $u_{\Delta}$. It must be explicit
(or more visible).
\\ \textbf{Response:} We compute the reference solution with a finely resolved collocation method. The one-dimensional testcase uses 100 Gauss-Legendre quadrature points (as mentioned in line 506 in the original manuscript), the two-dimensional testcase uses $50^2$ quadrature points (as mentioned in line 620 in the original manuscript) and the three-dimensional testcase uses $50^3$ quadrature points (as mentioned in line 661 in the original manuscript). You are correct that this information gets lost in the text which is why we added a short description whenever we plot reference solutions (which we do in Figures 1, 7 and 8).
\item Figure 3: the threshold $\varepsilon = 10^{-6}$ is not visible on the plots, as it should be. Also the intrusive and non intrusives methods must be more explicit: which ones (number of equations, . . . ).
\\ \textbf{Response:} We added the names of intrusive and non-intrusive methods and include references to the corresponding algorithms. The residual cannot be depicted in the plot itself, since the y-axis is the relative error, not the residual. We added references to the relative error as well as to the residual equations to avoid confusion.

@@

\item Lines 553/554 page 17 seem weird. What do you call positivity of mass?
Is it total mass? If yes, there is not problem with a conservative scheme.
\\ \textbf{Response:} We meant to say density, which we changed in the manuscript. The total mass (here the expected value of the total mass) is conserved by the scheme, however the solution at individual values for $\xi$ might yield negative densities, since we are using polynomials to span the solution in $\xi$. 

### 2020-05-19 Paper1/results.tex
review done?

\begin{figure}[h!]
\centering
	\begin{subfigure}{0.329\linewidth}
		\centering
				\includegraphics[width=\linewidth]{figs/Euler1DPlots5/sc100_ERho.png}
		\label{fig:referenceSolutionsub1}
	\end{subfigure}
	\hfill
	\begin{subfigure}{0.329\linewidth}
		\centering
				\includegraphics[width=\linewidth]{figs/Euler1DPlots5/sc100_VarRho.png}
		\label{fig:referenceSolutionsub2}
	\end{subfigure}
	\hfill
	\begin{subfigure}{0.329\linewidth}
		\centering
				\includegraphics[width=\linewidth]{figs/Euler1DPlots5/naca_mesh.png}
		\label{fig:referenceSolutionsMesh}
	\end{subfigure}
	\caption{Reference solution E$[\rho]$ and Var$[\rho]$ and the mesh close to the airfoil which is used in the computation of all presented methods.}
	\label{fig:referenceSolution}
\end{figure}

The aim is to quantify the effects arising from the one-dimensional uncertainty $\xi$ and to investigate its effects on the solution with different methods. To be able to measure the quality of the obtained solutions, we compute a reference solution using Stochastic Collocation with $100$ Gauss-Legendre quadrature points, which can be found in Figure~\ref{fig:referenceSolution}. In the following, we investigate the L$^2$-error of the variance and the expectation value. The L$^2$-error of the discrete quantity $\bm e_{\Delta}=(\bm e_1,\cdots,\bm e_{N_x})^T$, where $\bm e_j$ is the cell average of the quantity $\bm e$ in spatial cell $j$, is denoted by
\begin{linenomath*}\begin{align*}
\Vert \bm e_{\Delta} \Vert_{\Delta} := \sqrt{\sum_{j=1}^{N_x} \Delta x_j \bm e_j^2}\;.
\end{align*}\end{linenomath*}
Hence, when denoting the reference solution by $\bm u_{\Delta}$ and the moments obtained with the numerical method by $\bm{\hat u}_{\Delta}$, we investigate the relative error
\begin{linenomath*}\begin{align*}
\frac{\Vert \text{E}[\bm u_{\Delta}] - \text{E}[\mathcal{U}(\bm{\hat u}_{\Delta})] \Vert_{\Delta}}{\Vert \text{E}[\bm u_{\Delta}] \Vert_{\Delta}} \qquad \text{ and }\qquad \frac{\Vert \text{Var}[\bm u_{\Delta}] - \text{Var}[\mathcal{U}(\bm{\hat u}_{\Delta})] \Vert_{\Delta}}{\Vert \text{Var}[\bm u_{\Delta}] \Vert_{\Delta}}.
\end{align*}\end{linenomath*}
The error is computed inside a box of one meter height and 1.1 meters length around the airfoil to prevent small fluctuations in the coarsely discretized far field from affecting the error.

@@

	\begin{subfigure}{0.5\linewidth}
		\centering
				\includegraphics[width=\linewidth]{figs/errorEuler/L2_error_E[rho]_sc.pdf}
		\caption{}
		\label{fig:sub33}
	\end{subfigure}%
	\begin{subfigure}{0.5\linewidth}
		\centering
				\includegraphics[width=\linewidth]{figs/errorEuler/L2_error_Var[rho]_sc.pdf}
		\caption{}
		\label{fig:sub34}
	\end{subfigure}
	\caption{Comparison of the relative L$^2$-error for the density for IPM related methods (first row) and of the best performing IPM method in comparison with SC (second row). All intrusive methods converge to a residual $\varepsilon=6\cdot 10^{-6}$, whereas all non-intrusive methods converge to a residual $\varepsilon=1\cdot 10^{-7}$. All computations are performed with 5 MPI threads.}
	\label{fig:L2ErrorSolution}
\end{figure}
\begin{figure}[h!]
\centering
	\begin{subfigure}{0.329\linewidth}
		\centering
		\includegraphics[width=\linewidth]{figs/Euler1DPlots5/EulerSC5-2Res1e-6_ERho.png}
		\label{fig:sub1}
	\end{subfigure}%
	\hfill
	\begin{subfigure}{0.329\linewidth}
		\centering
		\includegraphics[width=\linewidth]{figs/Euler1DPlots5/sg_4_9_ERho.png}
		\label{fig:sub2}
	\end{subfigure}%
	\hfill
	\begin{subfigure}{0.329\linewidth}
		\centering
		\includegraphics[width=\linewidth]{figs/Euler1DPlots5/osIPM4-4_ERho.png}
		\label{fig:sub1}
	\end{subfigure}\\
	\vspace{-0.35cm}
	\begin{subfigure}{0.329\linewidth}
		\centering
		\includegraphics[width=\linewidth]{figs/Euler1DPlots5/EulerSC5-2Res1e-6_VarRho.png}
		\label{fig:sub1}
	\end{subfigure}%
	\hfill
	\begin{subfigure}{0.329\linewidth}
		\centering
		\includegraphics[width=\linewidth]{figs/Euler1DPlots5/sg_4_9_VarRho.png}
		\label{fig:sub2}
	\end{subfigure}%
	\hfill
	\begin{subfigure}{0.329\linewidth}
		\centering
		\includegraphics[width=\linewidth]{figs/Euler1DPlots5/osIPM4-4_VarRho.png}
		\label{fig:sub1}
	\end{subfigure}
	\caption{E$[\rho]$ and Var$[\rho]$ computed with SC$^5$, SG$_4$, IPM$_4$ (from left to right). Compare to the reference solution shown in Figure \ref{fig:referenceSolution}.}
	\label{fig:ERhoVarRho5}
\end{figure}

@@

First, let us mention that the adaptive SG method fails, since it yields negative densities during the iteration. The standard SG method however preserves positivity of mass, energy and pressure. The change of the relative L$^2$-error during the iteration to the steady state has been recorded in Figure~\ref{fig:sub31} for the expectation value and in Figure~\ref{fig:sub32} for the variance. When comparing intrusive methods without acceleration techniques as well as SC, the following properties emerge:
\begin{itemize}
\item Compared to IPM, stochastic-Galerkin comes at a significantly reduced runtime, meaning that the IPM optimization problem requires a significant computational effort.
\item For the expectation value SG$_9$ shows a smaller error compared to IPM$_9$, while for the variance, we see the opposite, i.e. IPM yields a better solution approximation than SG.
\end{itemize}
The proposed acceleration techniques show the following behavior:
\begin{itemize}
\item The One-Shot IPM (osIPM) method proposed in Section~\ref{sec:OneShotIPM} reduces the runtime while yielding the same error as the classical IPM method.
\item When using adaptivity (see Algorithm~\ref{alg:ad-IPM}) in combination with the One-Shot idea, the method is denoted by \textit{adaptive One-Shot IPM} (adosIPM). This method reaches the steady state IPM solution with a faster runtime than SG.
\item The idea of refinement retardation combined with adosIPM (see Algorithm~\ref{alg:adosIPM}) is denoted by \textit{retardation adosIPM} (readosIPM), which further decreases runtime. Here, we use two different strategies to increase the accuracy: First, we steadily increase the maximal truncation order when the residual approaches zero. To determine residual values for a given set of truncation orders $2,4,5$ and $8$, we study at which residual level the IPM method reaches a saturated error level for each truncation order. The residual values are then determined to be $6\cdot 10^{-5},3\cdot 10^{-5},2.2\cdot 10^{-5}$ and $2\cdot 10^{-5}$. The second, straight forward strategy converges the solution on a low truncation order of $2$ to a residual of $10^{-5}$ and then switches to a maximal truncation order of $9$. Strategy 1 is depicted in purple, Strategy 2 is depicted in yellow. It can be seen that both approaches reach the IPM$_9$ error for the same run time. Hence, we deduce that a naive choice of the refinement retardation strategy suffices to yield a satisfactory behavior.
\end{itemize}

@@

	\caption{Reference solution (\subref{fig:referenceSolutionsub2DE}) E$[\rho]$ and (\subref{fig:referenceSolutionsub2DVar}) Var$[\rho]$ and with logarithmic scaled (\subref{fig:referenceSolutionsub2DVarLog}) Var$[\rho]$. readosIPM$_{1-9}$ solution for (\subref{fig:reados2DE}) E$[\rho]$ and (\subref{fig:reados2DVar}) Var$[\rho]$, with the resulting (\subref{fig:readosIPMEVar2DRefinement}) refinement levels.}

@@

	\caption{Reference solution E$[\rho]$ and Var$[\rho]$ and used computational mesh (top row) and corresponding adaptive IPM solutions for sparse (middle row) and tensorized grids (bottom row) with the used refinement levels at the last time step}

>>>>

\begin{figure}[h!]
\centering
	\begin{subfigure}{0.329\linewidth}
		\centering
				\includegraphics[width=\linewidth]{figs/Euler1DPlots5/sc100_ERho.png}
		\label{fig:referenceSolutionsub1}
	\end{subfigure}
	\hfill
	\begin{subfigure}{0.329\linewidth}
		\centering
				\includegraphics[width=\linewidth]{figs/Euler1DPlots5/sc100_VarRho.png}
		\label{fig:referenceSolutionsub2}
	\end{subfigure}
	\hfill
	\begin{subfigure}{0.329\linewidth}
		\centering
				\includegraphics[width=\linewidth]{figs/Euler1DPlots5/naca_mesh.png}
		\label{fig:referenceSolutionsMesh}
	\end{subfigure}
	\caption{Reference solution E$[\rho]$ and Var$[\rho]$ and the mesh close to the airfoil which is used in the computation of all presented methods. \comment{The reference solution has been computed with stochastic-Collocation using $100$ Gauss-Legendre quadrature points.}}
	\label{fig:referenceSolution}
\end{figure}

The aim is to quantify the effects arising from the one-dimensional uncertainty $\xi$ and to investigate its effects on the solution with different methods. To be able to measure the quality of the obtained solutions, we compute a reference solution using Stochastic Collocation with $100$ Gauss-Legendre quadrature points, which can be found in Figure~\ref{fig:referenceSolution}. In the following, we investigate the L$^2$-error of the variance and the expectation value. The L$^2$-error of the discrete quantity $\bm e_{\Delta}=(\bm e_1,\cdots,\bm e_{N_x})^T$, where $\bm e_j$ is the cell average of the quantity $\bm e$ in spatial cell $j$, is denoted by
\begin{linenomath*}\begin{align*}
\Vert \bm e_{\Delta} \Vert_{\Delta} := \sqrt{\sum_{j=1}^{N_x} \Delta x_j \bm e_j^2}\;.
\end{align*}\end{linenomath*}
Hence, when denoting the reference solution by $\bm u_{\Delta}$ and the moments obtained with the numerical method by $\bm{\hat u}_{\Delta}$, we investigate the relative error
\begin{linenomath*}\begin{align}\label{eq:relErrorEVar}
\frac{\Vert \text{E}[\bm u_{\Delta}] - \text{E}[\mathcal{U}(\bm{\hat u}_{\Delta})] \Vert_{\Delta}}{\Vert \text{E}[\bm u_{\Delta}] \Vert_{\Delta}} \qquad \text{ and }\qquad \frac{\Vert \text{Var}[\bm u_{\Delta}] - \text{Var}[\mathcal{U}(\bm{\hat u}_{\Delta})] \Vert_{\Delta}}{\Vert \text{Var}[\bm u_{\Delta}] \Vert_{\Delta}}.
\end{align}\end{linenomath*}
The error is computed inside a box of one meter height and 1.1 meters length around the airfoil to prevent small fluctuations in the coarsely discretized far field from affecting the error.

@@

	\begin{subfigure}{0.5\linewidth}
		\centering
				\includegraphics[width=\linewidth]{figs/errorEuler/L2_error_E[rho]_sc.pdf}
		\caption{}
		\label{fig:sub33}
	\end{subfigure}%
	\begin{subfigure}{0.5\linewidth}
		\centering
				\includegraphics[width=\linewidth]{figs/errorEuler/L2_error_Var[rho]_sc.pdf}
		\caption{}
		\label{fig:sub34}
	\end{subfigure}
	\caption{Comparison of the relative L$^2$-error \comment{\eqref{eq:relErrorEVar}} for the density for IPM related methods (first row) and of the best performing IPM method in comparison with SC (second row). All intrusive methods \comment{(i.e. SG, IPM, osIPM, adosIPM, readosIPM) are converged} to a residual $\varepsilon=6\cdot 10^{-6}$ \comment{(see equation \eqref{eq:residualSteady})}, whereas all non-intrusive methods \comment{(i.e. SC) are converged} to a residual $\varepsilon=1\cdot 10^{-7}$ \comment{in order to obtain an error which remains constant in time. SG and IPM are computed with Algorithm~\ref{alg:IPM} (where SG uses a quadratic entropy) adapted for steady problems. For osIPM we used Algorithm~\ref{alg:osIPM}, for adosIPM and readosIPM we use Algorithm~\ref{alg:adosIPM} with and without refinement retardation.} All computations are performed with 5 MPI threads.}
	\label{fig:L2ErrorSolution}
\end{figure}
\begin{figure}[h!]
\centering
	\begin{subfigure}{0.329\linewidth}
		\centering
		\includegraphics[width=\linewidth]{figs/Euler1DPlots5/EulerSC5-2Res1e-6_ERho.png}
		\label{fig:sub1}
	\end{subfigure}%
	\hfill
	\begin{subfigure}{0.329\linewidth}
		\centering
		\includegraphics[width=\linewidth]{figs/Euler1DPlots5/sg_4_9_ERho.png}
		\label{fig:sub2}
	\end{subfigure}%
	\hfill
	\begin{subfigure}{0.329\linewidth}
		\centering
		\includegraphics[width=\linewidth]{figs/Euler1DPlots5/osIPM4-4_ERho.png}
		\label{fig:sub1}
	\end{subfigure}\\
	\vspace{-0.35cm}
	\begin{subfigure}{0.329\linewidth}
		\centering
		\includegraphics[width=\linewidth]{figs/Euler1DPlots5/EulerSC5-2Res1e-6_VarRho.png}
		\label{fig:sub1}
	\end{subfigure}%
	\hfill
	\begin{subfigure}{0.329\linewidth}
		\centering
		\includegraphics[width=\linewidth]{figs/Euler1DPlots5/sg_4_9_VarRho.png}
		\label{fig:sub2}
	\end{subfigure}%
	\hfill
	\begin{subfigure}{0.329\linewidth}
		\centering
		\includegraphics[width=\linewidth]{figs/Euler1DPlots5/osIPM4-4_VarRho.png}
		\label{fig:sub1}
	\end{subfigure}
	\caption{E$[\rho]$ and Var$[\rho]$ computed with SC$^5$, SG$_4$, IPM$_4$ (from left to right). Compare to the reference solution shown in Figure \ref{fig:referenceSolution}.}
	\label{fig:ERhoVarRho5}
\end{figure}

@@

First, let us mention that the adaptive SG method fails, since it yields negative \comment{gas} densities during the iteration. The standard SG method \comment{without adaptivity} however preserves positivity of \comment{density}, energy and pressure \comment{in our numerical experiments. Note that in general, stochastic-Galerkin does not guarantee positivity of these quantities, since the polynomial approximation does not guarantee any solution bounds.} The change of the relative L$^2$-error during the iteration to the steady state has been recorded in Figure~\ref{fig:sub31} for the expectation value and in Figure~\ref{fig:sub32} for the variance. When comparing intrusive methods without acceleration techniques as well as SC, the following properties emerge:
\begin{itemize}
\item Compared to IPM, stochastic-Galerkin comes at a significantly reduced runtime, meaning that the IPM optimization problem requires a significant computational effort.
\item For the expectation value SG$_9$ shows a smaller error compared to IPM$_9$, while for the variance, we see the opposite, i.e. IPM yields a better solution approximation than SG.
\end{itemize}
The proposed acceleration techniques show the following behavior:
\begin{itemize}
\item The One-Shot IPM (osIPM) method proposed in Section~\ref{sec:OneShotIPM} reduces the runtime while yielding the same error as the classical IPM method.
\item When using adaptivity (see Algorithm~\ref{alg:ad-IPM}) in combination with the One-Shot idea, the method is denoted by \textit{adaptive One-Shot IPM} (adosIPM). This method reaches the steady state IPM solution with a faster runtime than SG.
\item The idea of refinement retardation combined with adosIPM (see Algorithm~\ref{alg:adosIPM}) is denoted by \textit{retardation adosIPM} (readosIPM), which further decreases runtime. Here, we use two different strategies to increase the accuracy: First, we steadily increase the maximal truncation order when the residual approaches zero. To determine residual values for a given set of truncation orders $2,4,5$ and $8$, we study at which residual level the IPM method reaches a saturated error level for each truncation order. The residual values are then determined to be $6\cdot 10^{-5},3\cdot 10^{-5},2.2\cdot 10^{-5}$ and $2\cdot 10^{-5}$. The second, straight forward strategy converges the solution on a low truncation order of $2$ to a residual of $10^{-5}$ and then switches to a maximal truncation order of $9$. Strategy 1 is depicted in purple, Strategy 2 is depicted in yellow. It can be seen that both approaches reach the IPM$_9$ error for the same run time. Hence, we deduce that a naive choice of the refinement retardation strategy suffices to yield a satisfactory behavior.
\end{itemize}

@@

	\caption{Reference solution (\subref{fig:referenceSolutionsub2DE}) E$[\rho]$ and (\subref{fig:referenceSolutionsub2DVar}) Var$[\rho]$ and with logarithmic scaled (\subref{fig:referenceSolutionsub2DVarLog}) Var$[\rho]$. \comment{The reference solution has been computed with stochastic-Collocation using $50^2$ Gauss-Legendre quadrature points.} readosIPM$_{1-9}$ solution for (\subref{fig:reados2DE}) E$[\rho]$ and (\subref{fig:reados2DVar}) Var$[\rho]$, with the resulting (\subref{fig:readosIPMEVar2DRefinement}) refinement levels.}

@@

	\caption{Reference solution E$[\rho]$ and Var$[\rho]$ and used computational mesh (top row) and corresponding adaptive IPM solutions for sparse (middle row) and tensorized grids (bottom row) with the used refinement levels at the last time step. \comment{The reference solution has been computed with stochastic-Collocation using $50^3$ Gauss-Legendre quadrature points.}}
